Delta Smelt Enclosure Study 2019 Hybrid Paper Diet Analysis

### 2023-06-09 smelt_2019_allseasons/Zoop 2019.Rmd
catching up on my git hub

---
librarytitle: "Zoop Plotting"
output: html_document
editor_options: 
  chunk_output_type: console
---
# load packages
```{r}
library(tidyverse)
library(lubridate)
library(Polychrome)
library(zooper)
```

@@

#bring in tow data
aug <- read_csv("smelt_2019_allseasons/data_raw/Field_Environmental_Data_Aug_qaqc'ed_11.04.20.csv")
oct <- read_csv("smelt_2019_allseasons/data_raw/Field_Environmental_Data_Oct_qaqc'ed_12.11.20.csv")
aug <- read_csv("data_raw/Field_Environmental_Data_Aug_qaqc'ed_11.04.20.csv")
oct <- read_csv("data_raw/Field_Environmental_Data_Oct_qaqc'ed_12.11.20.csv")

@@

#calculate abundance -----------------------
all$distance <- (all$Revs)*57560/999999 #57,560 = low speed rotor constant
all$volumesampled <- ((pi*(all$RingSize_m^2))/4)*all$distance
all$samplecount <- (all$TotalVolume_ml/all$Subsampled_ml)*all$Count

@@

winter <- read_csv("smelt_2019_allseasons/data_clean/zoop_abundance.csv")
winter <- read_csv("data_clean/zoop_abundance.csv")

@@

library(vegan)
source("PlotNMDS.R")

@@

#first let's look at Rio Vista across seasons
RVzoops = filter(zoops, Location == "RV")

#raw abundance matrix
RVzoopmat = as.matrix(RVzoops[,6:12])

#relative abundance matrix
RVzoopmat2 = RVzoopmat/rowSums(RVzoopmat)

@@

RVb = adonis2(RVzoopmat2~Season, data = RVzoops)
RVb

@@

RVb = adonis2(RVzoopmat2~Season, data = RVzoops)
RVb

```
pretty big differences between seasons, R2 of 54% on the absolute abundance, 64% on relative abundance

THough I may need to update this with some sort of variable to account for day of year. Not sure.

@@

Fallb = adonis2(Fallzoopmat2~Location, data = Fallzoops)
Fallb
#

@@

N1 = metaMDS(zoopmat2)
zoops = mutate(zoops, Season  = as.factor(Season), Location = as.factor(Location))
PlotNMDS(N1, zoops, group = "Season")
PlotNMDS(N1, zoops, group = "Location")

RVN1 = metaMDS(RVzoopmat2)
RVzoops = mutate(RVzoops, Season  = as.factor(Season), Location = as.factor(Location))
PlotNMDS(RVN1, RVzoops, group = "Season")
PlotNMDS(RVN1, RVzoops, group = "Location")

@@

#remove new date column
weird.tag <- subset(weird.tag, select = -6)

@@

@@

zoopdiet = bind_rows(zoops, dietwide) %>%
  mutate(Amphipod = case_when(is.na(Amphipod) ~0, TRUE ~ Amphipod),
         Diptera = case_when(is.na(Diptera) ~0, TRUE ~ Diptera)) %>%
  select(-empty, -Microzoo)

@@

@@

RVda = adonis2(RVzoopdietmat2~ datatype + Season, data = RVdietzoops)
RVda

@@

ggplot(zoopdietlong, aes(x = datatype, y = count, fill = Taxa))+
  geom_col(position = "fill")+
  facet_grid(Location~Season)

>>>>

---
librarytitle: "Zoop Plotting"
output: html_document
editor_options: 
  chunk_output_type: console
---
# load packages
```{r}
library(tidyverse)
library(lubridate)
library(Polychrome)
library(zooper)
library(broom)
library(vegan)
```

@@

#bring in tow data
aug <- read_csv("smelt_2019_allseasons/data_raw/Field_Environmental_Data_Aug_qaqc'ed_11.04.20.csv")
oct <- read_csv("smelt_2019_allseasons/data_raw/Field_Environmental_Data_Oct_qaqc'ed_12.11.20.csv")
#aug <- read_csv("data_raw/Field_Environmental_Data_Aug_qaqc'ed_11.04.20.csv")
#oct <- read_csv("data_raw/Field_Environmental_Data_Oct_qaqc'ed_12.11.20.csv")

@@

#calculate abundance -----------------------
all$distance <- (all$Revs)*57560/999999 #57,560 = low speed rotor constant
all$volumesampled <- ((pi*(all$RingSize_m^2))/4)*all$distance

all$samplecount <- (all$TotalVolume_ml/all$Subsampled_ml)*all$Count

@@

winter <- read_csv("smelt_2019_allseasons/data_clean/zoop_abundance.csv")
#winter <- read_csv("data_clean/zoop_abundance.csv")

@@

source("PlotNMDS.R")

@@

#first let's look at Rio Vista across seasons
RVzoops = filter(zoops, Location == "RV")
fallzoops = filter(zoops, Season == "Fall")
#raw abundance matrix
RVzoopmat = as.matrix(RVzoops[,6:12])
fallzoopmat = as.matrix(fallzoops[,6:12])
#relative abundance matrix
RVzoopmat2 = RVzoopmat/rowSums(RVzoopmat)
fallzoopmat2 = fallzoopmat/rowSums(fallzoopmat)

@@

falla = adonis2(fallzoopmat2~Location*datatype, data = fallzoops)

##################################This is the one for the paper
RVb = adonis2(RVzoopmat2~Season, data = RVzoops)
RVb
RVresults = tidy(RVb)
write.csv(RVresults, file = "outputs/RVresults.csv")

@@

RVb = adonis2(RVzoopmat2BM~Season, data = RVzoops)
RVb

```
pretty big differences between seasons, R2 of 54% on the absolute abundance, 64% on relative abundance
pretty big differences between seasons, R2 of 65% on the absolute biomass, 95% on relative biomass
THough I may need to update this with some sort of variable to account for day of year. Not sure.

@@

######################################################
#this is the second permanova for the paper
Fallb = adonis2(Fallzoopmat2~Location, data = Fallzoops)
Fallb
write.csv(tidy(Fallb), "outputs/FallZoops.csv", row.names = F)

#

@@

N1 = metaMDS(zoopmat2, trymax = 50, try = 20)
zoops = mutate(zoops, Season  = as.factor(Season), Location = as.factor(Location))
PlotNMDS(N1, zoops, group = "Season")
PlotNMDS(N1, zoops, group = "Location")

RVN1 = metaMDS(RVzoopmat2)
RVzoops = mutate(RVzoops, Season  = as.factor(Season), Location = as.factor(Location))
PlotNMDS(RVN1, RVzoops, group = "Season")

@@

#remove new date column
weird.tag <- subset(weird.tag, select = -X.date)

@@

save(diet, file = "Dietdata.Rdata")

@@

```{r}
zoopdiet = bind_rows(zoops, dietwide) %>%
  mutate(Amphipod = case_when(is.na(Amphipod) ~0, TRUE ~ Amphipod),
         Diptera = case_when(is.na(Diptera) ~0, TRUE ~ Diptera)) %>%
  select(-empty, -Microzoo)

@@

#Now fall
Falldietzoops = filter(zoopdiet, Season == "Fall")

@@

Fallzoopdietmat = as.matrix(Falldietzoops[,c(6:11, 13, 14)])
Fallzoopdietmat2 = Fallzoopdietmat/rowSums(Fallzoopdietmat)

##########################This is the analysis for the paper!!!!
RVda = adonis2(RVzoopdietmat2~ datatype*Season, data = RVdietzoops)
RVda
write.csv(tidy(RVda), "outputs/DietZooppermanova.csv")

Fallda = adonis2(Fallzoopdietmat2~ Location*datatype, data = Falldietzoops)

Fallda
write.csv(tidy(Fallda), "outputs/DietZooppermanovaFall.csv")

Let's look at this a little more closely,s ee if we see any patterns.

```{r}
#It's really odd that data type isn't that important given some critters that show up in diet and not zooplankton at all.

ggplot(zoopdietlong, aes(x = SampleID, y = count, fill = Taxa)) + geom_col()+
  facet_wrap(~datatype, scales = "free")

ggplot(filter(zoopdietlong, Location == "RV"), aes(x = SampleID, y = count, fill = Taxa)) + geom_col()+
  facet_wrap(Season~datatype, scales = "free")

ggplot(filter(zoopdietlong, Location == "RV"), aes(x = SampleID, y = count, fill = Taxa)) + geom_col(position = "fill")+
  facet_wrap(Season~datatype, scales = "free")

I tested for an interaction between datatype and season, but nothing. I think it's just too variable to say much.

```{r}
#NMDS's that we probably won't include

@@

zoopnos = group_by(zoopdiet, datatype, Season, Location) %>%
  summarize(N = length(unique(SampleID)))

ggplot(zoopdietlong, aes(x = datatype, y = count, fill = Taxa))+
  geom_col(position = "fill")+
  facet_grid(Location~Season)+
  scale_fill_brewer(palette = "Dark2", name = "Taxon")+
  theme_bw()+
  ylab("Relative Abundance")+
  geom_label(data = zoopnos, aes(x = datatype, y = 1, label = N), inherit.aes = FALSE)+
  scale_x_discrete(labels = c("Zooplankton", "Diets"), name = NULL)

ggsave("outputs/zoopdiet.tiff", device = "tiff", width = 8, height = 6.5)
